---
title: "Class 1 Exercises"
output: html_notebook
---
##My first exercise

```{r}
classFirstVar <- "Shangri La"

classFirstVar
```


##Notes

Data Structures
Vector
1
2
3
4
5
6
7
10
132

Pet Name
c is short for combine, it takes what's in the parentheses and combines them into a vector.

```{r}
petNames <- c("Tommy", "Singha", "Chacko")
petAge <- c(4, 5, 14)
male <- c(TRUE, FALSE, FALSE)
```

```{r}
typeof(petNames)
typeof(petAge)
typeof(male)
```

```{r}

```


4. What happens when you combine different data types? 
Coercion hierarchy: logical, integer, double, character
anytime R encounters multiples kinds of ___ it will convert it to the lowest possible thing, which is character
Bullion
```{r}
c("Hi", 1, "one")
```

```{r}
c(1.0, 3.4, 5L)
```

```{r}
c(T, F, "TRUE")
```

```{r}
c(T, F, 1)
```

###Lists
1. Create a list

```{r}
testList <- (hiPets)
```

2. Create a named list and access its names using 

###Data Frame



1. Create and store a character vector (of your choice), a numeric vector (integer/double) and a factor, all of the same length
```{r}
plantNames <- c("orchid", "calla lily", "fern")
plantAbundance <- c(50, 60, 85)
```



2. Use `seq()` to generate an atomic vector containing all the odd numbers less than 100
```{r}
##Create a sequence of odd numbers from 1 to 99
seq(from = 1, to = 100, by = 2)
```


3. Store all the vectors created in step 1 in a named list
```{r}
plantList <- list(plantNames, plantAbundance)

plantList
```


4. Create a dataframe using all the vectors stored in step 1
```{r}
taskDF <- data.frame(plantNames, plantAbundance)

taskDF
```


5. Predict the output of the following without running the code in `c("FALSE", FALSE)`, `c(1.0, FALSE)`, `c(2, TRUE)` and `c("two", 2)`
  `c("FALSE", FALSE)` = "Character"
  `c(1.0, FALSE)` = "Numeric (Integer/Double)"
  `c(2, TRUE)` = "Numeric"
  `c("two", 2)` = "Character"

```{r}
c("FALSE", FALSE)
c(1.0, FALSE)
c(2, TRUE)
c("two", 2)
```


6. What gets returned when you check the `typeof()` the dataframe created in step 4. Why?

```{r}
installed.packages()
```



